The Need to Develop a Holistic Infectious Disease Surveillance System

### 2024-10-24 manuscript/manuscript.qmd
Add comments of things to update in manuscript

@@

@@

- No test, where every suspected measles case is counted as positive for measles and used to trigger an outbreak alert, which is equivalent to the use of the clinical case definition. This is implemented as a diagnostic test with 100% sensitivity, 0% specificity, a 0-day lag in returning the result, and all individuals who visit the clinic are tested (not varied between 10-60%, like the other diagnostic tests described).
- An RDT equivalent with 85% sensitivity and specificity, and 0-day lag in result return. That is, 85% of true measles cases will be correctly labelled as positive, and 15% of non-measles febrile rash individuals that are tested will be incorrectly labelled as positive for measles. This acts as a lower bound of acceptability for a new measles RDT.
- An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return.
- An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay.
- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.
- An ELISA equivalent test with 100% sensitivity and specificity with a 14-day test result delay.

@@

> Despite the same average incidence of noise in each (comparable) scenario, the relative proportion of measles to noise varies differently throughout the time series, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
Imperfect tests are also more susceptible to incorrect results because of the varying prevalence throughout the time series, so the PPV of the system will not be static throughout the time series.

Under the assumption that non-measles febrile rash is relatively static in time (Poisson), RDTs can perform as well, if not better than ELISA tests at moderate to high testing rates, and at a fraction of the cost.
However, if it is expected that the noise is non-stationary, imperfect tests cannot overcome their accuracy limitations through higher testing rates, saturating at c. 74% accuracy, relative to ELISA’s 93%.

@@

>>>>

_Mention that the same model was used to define the noise as the outbreaks._

@@

_Need to mention that an alert is any time the threshold is exceeded, regardless of the number of consecutive days of exceedance._
_Clearly state that 1 day of exceedance is and 30 days of consecutive exceedance are both treated as a single alert i.e., the same._

@@

<!-- - No test, where every suspected measles case is counted as positive for measles and used to trigger an outbreak alert, which is equivalent to the use of the clinical case definition. This is implemented as a diagnostic test with 100% sensitivity, 0% specificity, a 0-day lag in returning the result, and all individuals who visit the clinic are tested (not varied between 10-60%, like the other diagnostic tests described). -->
- An RDT equivalent with 85% sensitivity and specificity, and 0-day lag in result return. That is, 85% of true measles cases will be correctly labelled as positive, and 15% of non-measles febrile rash individuals that are tested will be incorrectly labelled as positive for measles. This acts as a lower bound of acceptability for a new measles RDT.
- An RDT equivalent with 90% sensitivity and specificity, and 0-day lag in result return.
- An ELISA equivalent test with 100% sensitivity and specificity with a 0-day test result delay.
<!--- An ELISA equivalent test with 100% sensitivity and specificity with a 7-day test result delay.-->
- An ELISA equivalent test with 100% sensitivity and specificity with a 14-day test result delay.

@@

Under the assumption that non-measles febrile rash is relatively static in time (Poisson), RDTs can perform as well, if not better than ELISA tests at moderate to high testing rates, and at a fraction of the cost.
However, if it is expected that the noise is non-stationary, imperfect tests cannot overcome their accuracy limitations through higher testing rates, saturating at c. 74% accuracy, relative to ELISA’s 93%.
This occurs because, despite the same average incidence of noise in each (comparable) scenario, the relative proportion of measles to noise varies differently throughout the time series, exacerbating the effects of imperfect diagnostic tests that will produce higher rates of false positives and negatives than ELISA-like diagnostics.
Imperfect tests are also more susceptible to incorrect results due to the varying prevalence throughout the time series, so the PPV of the system will not be static throughout the time series.

- _Make point about balance of false positive/negative outbreak detection_
- _Analysis isn't true optimization:_
  - _Requires explicit decisions about preference for speed / false alerts vs higher PPV_
  - _Surveillance is counting for action (WHO quote)_

@@

- _Something about GAVI/Gates_

### 2024-10-24 manuscript/supplemental-appendix.qmd
Add comments of things to update in manuscript

The outbreak detection accuracy for each testing scenario, at the optimal thresholds.

@@

Unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

@@

Outbreak alert delay (days) of each testing scenario at their specific optimal thresholds. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

>>>>

Mean outbreak detection accuracy of each testing scenario at their specific optimal thresholds, when the average noise incidence is 8 times higher than the average measles incidence. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

@@

Mean unavoidable cases per annum of each testing scenario at their specific optimal thresholds, scaled up to Ghana’s 2022 population, when the average noise incidence is 8 times higher than the average measles incidence. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.

@@

Mean outbreak alert delay (days) of each testing scenario at their specific optimal thresholds, when the average noise incidence is 8 times higher than the average measles incidence. A) the noise structure is dynamical, and the seasonality is in-phase with the measles incidence. B) the noise structure is Poisson only.
